---
title: "Day 2 distributions"
output:
  pdf_document: default
  html_notebook: default
---
## Binomial Distribution 

In the experiment where there are only two out come of the result. An example getting the value 6 in ludo. 

Question: What is the probability of getting 6 in ludo.
P=1/6
q=1-p
q=5/6

Now given the probability of getting 6 in ludo is P, if you though how many 6 you get in the game. In the game you played 50 times. 

Question : Given that you played game 6 time what is the probability that you get 5 sixes. This can be found using binomial distribution function. 

$P(X = k) = \binom{n}{k} p^k (1-p)^{n-k}$

lets solve this. 

We can also used r function can dbinom(success, size=20, prob=.3)
```{r dbinom, echo=TRUE}
```

## Normal Distribution



$P(x) = \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right. \kern-\nulldelimiterspace} {2\sigma ^2 }}}$

rnorm(100000,1,5),main = "",xlab = "")

## Poisson Distribution

$P\left( x \right) = \frac{{e^{ - \lambda } \lambda ^x }}{{x!}}$

rpois(100000,4),main = "",xlab = " ")

## Anova

$t = \mu{1}-\mu{2}/\sqrt(\sigma^2*(1/N{1}+1/N{2})$

pooled S


$\sigma^2=((N{1}-1)*s1^2+(N{2}-1)*s2^2)/(N{1}+N{2}-2)$

```{r anova, echo=T}
teenage_birth <- read.csv("./data/Percentage_teen_birth.csv",header = T)
model1<-aov(value~State.group,teenage_birth)
summary(model1)
```