---
title: "Linear regression model"
author: "Chandan Kumar Pandey"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

# Linear regression with only one predictor. 

### Student marks dataset and exploration. 

The data set contain the mark scored by student in exam, hours of study and number of course they have opted for. The question we are asking in this case if, __How does number of hours of studies will impact score? __

```{r LM simple not transformation, echo=T,warning=F}
library(ggplot2)
library(GGally)
#Read the data set
Marks <- read.csv("data/Student_Marks.csv",header = T)
# Basic exploration of the marks data set. 
summary(Marks)
#graphical representations. 
boxplot(Marks$Marks)
boxplot(Marks$time_study)
str(Marks)
# The first set of data exploration is pair plot which check for 
#correlation among predictor and repose.
ggpairs(Marks[,c(3,2,1)])
```

### Linear regression fitting. 

Based on Pairwise plot it is clear that Marks do follow a linear trend with time of study. In the next step we will model the linear regression

$$y = \beta{0} + \beta{1}X+\epsilon$$ where $$\epsilon= N(\mu,\sigma)$$

```{r LM fiting, echo=T}
model <- lm(Marks~time_study,data = Marks)
#model's assumptions validation. 
par(mfrow=c(2,2))
plot(model)
```

### Assumption of linear regression. 

1.  Linear relation.
2.  Normality
2.  Homogeneity of residuals variance

Based on out graphs above, the plot of Residuals vs Fitted value we can see that out data in not homogeneous. We can later confirm this by ploting histogram of Marks.
__One way to overcome this problem is to do _transformation_.__

```{r ,echo=T}
library(ggplot2)
ggplot(data = Marks,aes(x=Marks))+geom_histogram(binwidth = 5)
##Log transformation
ggplot(data = Marks,aes(x=log(Marks)))+geom_histogram(binwidth = 0.5)
##lets check the normality by using shapiro test
shapiro.test(Marks$Marks)
shapiro.test(log(Marks$Marks))
## The value p-value have increase but not to significant level that 
## distribution become normal.
```

The value p-value have increase but not to significant level that 
distribution become normal. However, given with data the model will improve significantly

```{r model with log transform, echo=T}
model1 <- lm(log(Marks)~time_study,data = Marks)
#model's assumptions validation. 
par(mfrow=c(2,2))
plot(model1)
summary(model1)
```

### interpretation of model with log transformation. 

$$log(Marks) = 1.978 + 0.253*study time + \epsilon$$

```{r ploting the linear model result, echo=TRUE}
plot_model2 <- ggplot(data = Marks,aes(x=time_study,y=log(Marks)))
plot_model2+geom_point(pch=1,col="blue")+
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE)+
  theme_bw()+
  labs(x="Time of study",y="Marks (in Log scale)")+
  geom_text(mapping = aes(x=4,y=3.8,
                          label=paste("R^2=",round(summary(model1)$adj.r.squared,4))))
```

In order to back transform from log scale to linear scale we need to look in the use the formula 

$$Marks = \exp^{B{1}*time+B{0}}$$
```{r reverse transform, echo=T}
exp_pred<-exp(predict(model1,newdata = Marks))
plot(Marks$Marks~Marks$time_study)
points(exp_pred~Marks$time_study,col="red")
```

## Model with more that one parameter. 

$$y = \beta{0} + \beta{1}X{1}+\beta{2}X{2}+\epsilon$$ where $$\epsilon= N(\mu,\sigma)$$

Let us use the fish morphometric data and ask question is fish length can predict its weight. 

```{r load fish data, echo=T}
fish <- read.csv("data/Fish_morphometrics.csv",header = T)
# Summary and str of data
summary(fish)
str(fish)
# pairwise plot
ggpairs(fish[,3:7])
library(dplyr)
fish_group<-fish%>%group_by(Species)%>%summarise(count=n(),mean(Weight),mean(Length3),sd(Weight))
fish<-fish%>%mutate(avg_length = (Length1+Length2+Length3)/3)
hist(fish$Weight)

```

```{r model with 2 predictor,echo=T}
#Install.packages("devtools")
devtools::install_github("cardiomoon/ggiraphExtra")
library(ggiraphExtra)
require(ggiraph)
model_fish1 <- lm(Weight~avg_length+Species,data = fish)
fish <- fish[fish$Weight!=0,] ## wrong data point
model_log <- lm(log(Weight)~avg_length+Species,data = fish)
#we make log log transformation
fish <- fish%>%mutate(log_w = log(Weight),log_avg_len=log(avg_length))
model_log_log <- lm(log_w~log_avg_len+Species,data = fish)
par(mfrow=c(2,2))
plot(model_log_log)
summary.lm(model_log_log)
summary.aov(model_log_log)
model_log_log_aov <- aov(log_w~log_avg_len+Species,data = fish)
par(mfrow=c(1,1))
#posthoc 
Tuk <-TukeyHSD(model_log_log_aov, "Species",data = fish)
plot(Tuk)
#predicting the values
pred_value<-predict.lm(model_log_log,interval = "confidence")
pred_value <- as.data.frame(pred_value)
fish <- cbind(fish,pred_value)
## plotting the model 
plot <- ggplot(data = fish, aes(x=log_avg_len,y=log_w,col=Species))
plot+geom_point()+
  geom_line(aes(y=fit))+
  geom_ribbon(aes(ymin=lwr,ymax=upr),alpha=0.05)
```


