---
title: "Poisson regression"
author: "Chandan Kumar Pandey"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

# Poisson Distribution

$$P(X=k) = {{\lambda^{k}*e^{-\lambda}}/k!}$$
$$ \lambda = e^{\beta{0}+ (\beta{1}*X{1})} $$

Example for poisson distribution. 

1. Number of goal scores in match. 
2. Number of car passing by traffic light. 

basically, 

a. count data
b. Rate

# Poisson regression

## When to use

1. Response is either is count or rate. 
2. Whole number 
3. The data may contain a large number of data points for just a few values, thereby making the frequency distribution quite skewed. See for example above histogram.
4. The data may reflect the occurrence of a rare event
5. It can be assumed that there is a certain rate of occurrence of events $\lambda$

## Data

We will be work on data where math score and number of award is provide. 
We will fit simple poisson model with one variable. 

```{r maths score, echo=T}
library(dplyr)
library(ggplot2)
maths <- read.csv("data/competition_awards_data.csv")
mod1 <- glm(Awards~Math.Score,data = maths,family = "poisson")
summary(mod1)
library(AER)
dispersiontest(mod1)

##plotting the model 
predicted_value <- predict.glm(mod1,maths,type = "response",se.fit = T)
predicted_value <- predicted_value%>%as.data.frame()
df <- cbind(M=maths$Math.Score,predicted_value)
ggplot(data = df ,aes(x=M,y=fit))+geom_line()+
  geom_ribbon(aes(ymin=fit-se.fit,ymax=fit+se.fit),alpha=0.5)+
  geom_point(data = maths,aes(x=Math.Score,y=Awards))
```

 In real life it is very hard to get data which follow this trend, therefore we use negative binomial distribution to model where overdisperssion is higher that 1.10.
 
## Example 

Road kill data
 
```{r load the data, echo=T,warning=F}
Road_kill <- read.csv("./data/road_kill.csv",header = T)
## summary()
summary(Road_kill)
library(dplyr)
Road_kill_sum <- Road_kill%>%group_by(Class,Year,Month)%>%
  summarise(count=n())%>%as.data.frame()
Road_kill_sum<-na.omit(Road_kill_sum)
class(Road_kill_sum$count)
Road_kill_sum$Month <- as.factor(Road_kill_sum$Month)

```

## If the number of road kill have increase with years


```{r model with poisson,echo=TRUE}
mod1 <- glm(count~Year,data = Road_kill_sum,family = "poisson")
summary(mod1)
## check for over dispersion 
library(AER)
dispersiontest(mod1)
## it is wrong but lets predict it. 
value_pred <- predict(mod1,Road_kill_sum,type = "response",se.fit = T)
value_pred <- as.data.frame(value_pred)
value_pred <- cbind(Years = Road_kill_sum$Year,value_pred)

```

Since th dispersion is very high we will used negative binomial 

# Negetive binomial distribution

```{r nb dis,echo=TRUE}
library(MASS)
mod_kill_nb <- glm.nb(count ~ Year, data = Road_kill_sum)
summary(mod_kill_nb)
value_nb_predict <- predict(mod_kill_nb,Road_kill_sum,type = "response",se.fit = T)
value_nb_predict <- as.data.frame(value_nb_predict)
value_nb_predict <- cbind(y1 = Road_kill_sum$Year,value_nb_predict)
#let plot wrong poisson model, nb model with data points
plot <- ggplot(data =value_pred,aes(x=Years,y=fit))
p_wrog <-plot+ geom_line()+
  geom_ribbon(aes(ymax=fit+se.fit,ymin=fit-se.fit))+
  geom_point(data = Road_kill_sum,aes(x=Year,y=count))
  
  

plot1 <- ggplot(data =value_nb_predict,aes(x=y1,y=fit))
P1 <-plot1+ geom_line()+
  geom_ribbon(aes(ymax=fit+se.fit,ymin=fit-se.fit))+
  geom_point(data = Road_kill_sum,aes(x=Year,y=count))
  
library(gggrid)
library(gridExtra)
grid.arrange(p_wrog,P1,nrow=2,ncol=1)
summary(mod_kill_nb)

mod_kill_nb_add <- glm.nb(count ~ Year+Class, data = Road_kill_sum)
mod_kill_nb_int <- glm.nb(count ~ Year*Class, data = Road_kill_sum)
anova(mod_kill_nb_int)
summary(mod_kill_nb_add)
summary(mod_kill_nb_int)
AIC(mod_kill_nb,mod_kill_nb_add,mod_kill_nb_int)
library(AICcmodavg)
#define list of models
models <- list(mod_kill_nb,mod_kill_nb_add,mod_kill_nb_int)

#specify model names
mod.names <- c('mod_kill_nb', 'mod_kill_nb_add',"mod_kill_nb_int")

#calculate AIC of each model
aictab(cand.set = models, modnames = mod.names)
```

```{r interactive model,echo=TRUE}
Year <- unique(Road_kill_sum$Year)
Class<-unique(Road_kill$Class)
Year <- c(Year,Year,Year,Year)
Class <- c(rep(Class[1],20),rep(Class[2],20),rep(Class[3],20),rep(Class[4],20))
newdata <- data.frame(Year,Class)
interactive_pred_value <- predict(mod_kill_nb_int,newdata, type = "response",se.fit = T)
interactive_pred_value <- as.data.frame(interactive_pred_value)
interactive_pred_value <- cbind(interactive_pred_value,newdata)
plot_int <- ggplot(data = interactive_pred_value,
                   aes(x=Year,y=fit,col=Class))
plot_int + geom_line()+ 
  geom_ribbon(aes(ymax=fit+se.fit,ymin=fit-se.fit,fill=Class),alpha=0.1)
```