---
title: "Review of Basic R "
author: "Chandan Kumar Pandey"
date: "`r Sys.Date()`"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Let us Revise

-   Variables
-   Operators
-   Objects
-   Conditional statements
-   Loops or iteration. 
-   Function
-   Bit of Plot

## What we will learn in next four days. 

### Day 2 {Friday}
-   Frequency distribution; Normal Distribution, Binomial distribution, and Poisson distribution. 
-   Linear model and plotting the difference between groups. 

## Examples

```{r cars, echo = F}
par(mfrow=c(2,2))
hist(rnorm(100000,1,5),main = "",xlab = "")
hist(rpois(100000,4),main = "",xlab = " ")
success <- 0:20
value <-dbinom(success, size=20, prob=.3)
plot(success,value ,type='h')
plot(1:100,1:100,xlab = "Number",ylab = "Number")
```

# Day3:
- Linear regression models; concept 
- Hands-on example of linear regression with a single predictor. 
- Hands-on example of linear regression with multiple predictors. 
- Plotting the linear models

# Day4:
- Generalized linear model, concept
- Hands-on example of GLM; Binomial distribution
- Hands-on example of GLM; Poisson distribution

# Day5: 
- Multiple Linear Regression with Interactions
- Model selection using AIC 

